---
title: "15 Insights from Washington School Enrollment Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{15 Insights from Washington School Enrollment Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 5,
  cache = TRUE
)
```

```{r load-packages}
library(waschooldata)
library(dplyr)
library(tidyr)
library(ggplot2)

theme_set(theme_minimal(base_size = 14))
```

This vignette explores Washington's public school enrollment data, surfacing key trends across 16 years of data (2010-2025).

---

## 1. Washington added 70,000 students over 16 years before COVID reversed the trend

The Evergreen State grew enrollment steadily from 1.03 million to nearly 1.15 million before the pandemic wiped out years of gains.

```{r statewide-data}
enr <- fetch_enr_multi(2010:2025, use_cache = TRUE)

state_totals <- enr |>
  filter(is_state, subgroup == "total_enrollment", grade_level == "TOTAL") |>
  select(end_year, n_students) |>
  mutate(change = n_students - lag(n_students),
         pct_change = round(change / lag(n_students) * 100, 2))
stopifnot(nrow(state_totals) > 0)

state_totals
```

```{r statewide-chart}
print(state_totals)

ggplot(state_totals, aes(x = end_year, y = n_students)) +
  geom_line(linewidth = 1.2, color = "#4B2E83") +
  geom_point(size = 3, color = "#4B2E83") +
  geom_vline(xintercept = 2020.5, linetype = "dashed", color = "red", alpha = 0.5) +
  annotate("text", x = 2020.8, y = max(state_totals$n_students, na.rm = TRUE),
           label = "COVID", hjust = 0, color = "red", size = 3) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Washington Public School Enrollment (2010-2025)",
    subtitle = "A decade of growth disrupted by the pandemic",
    x = "School Year (ending)",
    y = "Total Enrollment"
  )
```

---

## 2. Seattle peaked at 56,000 students then lost 9% in five years

Washington's largest district peaked in 2020 and has lost nearly 5,000 students since, even as tech industry growth transformed the region.

```{r top-districts-data}
seattle <- enr |>
  filter(is_district, subgroup == "total_enrollment", grade_level == "TOTAL",
         grepl("Seattle", district_name, ignore.case = TRUE)) |>
  select(end_year, district_name, n_students) |>
  mutate(pct_of_peak = round(n_students / max(n_students) * 100, 1))
stopifnot(nrow(seattle) > 0)

seattle
```

```{r top-districts-chart}
top_districts <- enr |>
  filter(is_district, subgroup == "total_enrollment", grade_level == "TOTAL",
         end_year == 2025) |>
  arrange(desc(n_students)) |>
  head(5) |>
  pull(district_id)
stopifnot(length(top_districts) > 0)

top5_data <- enr |>
  filter(is_district, subgroup == "total_enrollment", grade_level == "TOTAL",
         district_id %in% top_districts)
stopifnot(nrow(top5_data) > 0)

print(top5_data |> filter(end_year == 2025) |> select(district_name, n_students))

ggplot(top5_data, aes(x = end_year, y = n_students, color = district_name)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  geom_vline(xintercept = 2020.5, linetype = "dashed", color = "red", alpha = 0.5) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Washington's Top 5 Districts: Enrollment Trends",
    subtitle = "Seattle peaked in 2020 while suburban districts continue growing",
    x = "School Year",
    y = "Enrollment",
    color = "District"
  ) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(nrow = 2))
```

---

## 3. White students dropped from 64% to 48% as Hispanic enrollment nearly doubled

The state's demographics reflect Pacific Rim immigration and a growing Hispanic population that rose from 167K to 295K students.

```{r demographics-data}
demographics <- enr |>
  filter(is_state, grade_level == "TOTAL",
         subgroup %in% c("white", "hispanic", "asian", "black", "multiracial"),
         end_year %in% c(2010, 2015, 2020, 2025)) |>
  select(end_year, subgroup, n_students, pct) |>
  mutate(pct = round(pct * 100, 1))
stopifnot(nrow(demographics) > 0)

demographics |>
  pivot_wider(names_from = end_year, values_from = c(n_students, pct))
```

```{r demographics-chart}
demo_trend <- enr |>
  filter(is_state, grade_level == "TOTAL",
         subgroup %in% c("white", "hispanic", "asian", "black", "multiracial"))
stopifnot(nrow(demo_trend) > 0)

print(demo_trend |> filter(end_year == 2025) |> select(subgroup, n_students, pct))

ggplot(demo_trend, aes(x = end_year, y = pct * 100, color = subgroup)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_color_brewer(palette = "Set2",
                     labels = c("Asian", "Black", "Hispanic", "Multiracial", "White")) +
  labs(
    title = "Washington's Shifting Demographics (2010-2025)",
    subtitle = "White enrollment declining as Hispanic and multiracial populations grow",
    x = "School Year",
    y = "Percent of Students",
    color = "Race/Ethnicity"
  ) +
  theme(legend.position = "bottom")
```

---

## 4. Puget Sound holds 427K students -- nearly 4x Eastern Washington

The I-5 corridor dominates enrollment, but Eastern Washington's districts face different challenges.

```{r regional-data}
esd_enrollment <- enr |>
  filter(is_district, subgroup == "total_enrollment", grade_level == "TOTAL",
         end_year == 2025, !is.na(esd_name)) |>
  group_by(esd_name) |>
  summarize(
    districts = n(),
    students = sum(n_students, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(students))
stopifnot(nrow(esd_enrollment) > 0)

esd_enrollment
```

```{r regional-chart}
print(esd_enrollment)

esd_enrollment |>
  mutate(esd_short = gsub("Educational Service District|ESD", "ESD", esd_name)) |>
  mutate(esd_short = forcats::fct_reorder(esd_short, students)) |>
  ggplot(aes(x = students, y = esd_short, fill = students)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = scales::comma(students)), hjust = -0.1, size = 3) +
  scale_x_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.2))) +
  scale_fill_gradient(low = "#E8E3D3", high = "#4B2E83") +
  labs(
    title = "Washington Enrollment by Educational Service District",
    subtitle = "Puget Sound ESD 121 dominates with 427K students",
    x = "Total Students",
    y = NULL
  )
```

---

## 5. Suburban districts are booming: Sumner-Bonney Lake up 23% since 2015

The fastest-growing large districts are spread across the state, from Sumner-Bonney Lake south of Seattle to Central Valley near Spokane.

```{r growth-data}
growth_rates <- enr |>
  filter(is_district, subgroup == "total_enrollment", grade_level == "TOTAL",
         end_year %in% c(2015, 2025)) |>
  select(district_id, district_name, end_year, n_students) |>
  pivot_wider(names_from = end_year, values_from = n_students, names_prefix = "yr_") |>
  filter(yr_2015 > 5000) |>
  mutate(
    change = yr_2025 - yr_2015,
    pct_change = round((yr_2025 - yr_2015) / yr_2015 * 100, 1)
  ) |>
  arrange(desc(pct_change))
stopifnot(nrow(growth_rates) > 0)

head(growth_rates, 10)
```

```{r growth-chart}
growth_top10 <- head(growth_rates, 10)
print(growth_top10)

growth_top10 |>
  mutate(district_name = forcats::fct_reorder(district_name, pct_change)) |>
  ggplot(aes(x = pct_change, y = district_name, fill = pct_change > 0)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0("+", pct_change, "%")),
            hjust = -0.1, size = 3) +
  scale_x_continuous(expand = expansion(mult = c(0.1, 0.15))) +
  scale_fill_manual(values = c("TRUE" = "#4B2E83", "FALSE" = "#C4A055")) +
  labs(
    title = "Fastest Growing Large Districts (2015-2025)",
    subtitle = "Among districts with 5,000+ students in 2015",
    x = "Percent Change",
    y = NULL
  )
```

---

## 6. The kindergarten cliff is real: K enrollment down 14% from 2020 peak

Washington's kindergarten enrollment dropped during COVID and hasn't recovered, signaling smaller cohorts ahead.

```{r k-enrollment}
k_trend <- enr |>
  filter(is_state, subgroup == "total_enrollment", grade_level == "K") |>
  select(end_year, n_students) |>
  mutate(pct_of_peak = round(n_students / max(n_students) * 100, 1))
stopifnot(nrow(k_trend) > 0)

k_trend
```

```{r k-chart}
print(k_trend)

ggplot(k_trend, aes(x = end_year, y = n_students)) +
  geom_line(linewidth = 1.2, color = "#4B2E83") +
  geom_point(size = 3, color = "#4B2E83") +
  geom_vline(xintercept = 2020.5, linetype = "dashed", color = "red", alpha = 0.5) +
  annotate("text", x = 2020.8, y = max(k_trend$n_students, na.rm = TRUE),
           label = "COVID", hjust = 0, color = "red", size = 3) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Washington Kindergarten Enrollment (2010-2025)",
    subtitle = "COVID-era losses persist: 71K in 2025 vs 83K peak in 2020",
    x = "School Year (ending)",
    y = "Kindergarten Enrollment"
  )
```

---

## 7. Spokane anchors Eastern Washington with nearly 30,000 students

Spokane Public Schools serves 29,690 students, making it the largest district east of the Cascades and the state's second-largest overall.

```{r spokane-data}
spokane <- enr |>
  filter(is_district, subgroup == "total_enrollment", grade_level == "TOTAL",
         grepl("Spokane", district_name, ignore.case = TRUE),
         end_year == 2025) |>
  select(district_name, n_students) |>
  arrange(desc(n_students))
stopifnot(nrow(spokane) > 0)

spokane
```

```{r spokane-chart}
spokane_trend <- enr |>
  filter(is_district, subgroup == "total_enrollment", grade_level == "TOTAL",
         grepl("^Spokane School District$", district_name))
stopifnot(nrow(spokane_trend) > 0)

print(spokane_trend |> select(end_year, n_students))

ggplot(spokane_trend, aes(x = end_year, y = n_students)) +
  geom_line(linewidth = 1.2, color = "#4B2E83") +
  geom_point(size = 3, color = "#4B2E83") +
  geom_vline(xintercept = 2020.5, linetype = "dashed", color = "red", alpha = 0.5) +
  annotate("text", x = 2020.8, y = max(spokane_trend$n_students, na.rm = TRUE),
           label = "COVID", hjust = 0, color = "red", size = 3) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Spokane School District Enrollment (2010-2025)",
    subtitle = "Eastern Washington's anchor district holding near 30K",
    x = "School Year (ending)",
    y = "Total Enrollment"
  )
```

---

## 8. 130 districts have fewer than 500 students

Nearly 40% of Washington's districts serve fewer than 500 students, challenging their long-term viability.

```{r small-districts-data}
district_sizes <- enr |>
  filter(is_district, subgroup == "total_enrollment", grade_level == "TOTAL",
         end_year == 2025) |>
  mutate(size_bucket = case_when(
    n_students < 500 ~ "Small (<500)",
    n_students < 2000 ~ "Medium (500-2K)",
    n_students < 10000 ~ "Large (2K-10K)",
    TRUE ~ "Very Large (10K+)"
  )) |>
  count(size_bucket)
stopifnot(nrow(district_sizes) > 0)

district_sizes
```

```{r small-districts-chart}
print(district_sizes)

district_sizes |>
  mutate(size_bucket = factor(size_bucket,
    levels = c("Small (<500)", "Medium (500-2K)", "Large (2K-10K)", "Very Large (10K+)"))) |>
  ggplot(aes(x = size_bucket, y = n, fill = size_bucket)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = n), vjust = -0.5, size = 4) +
  scale_fill_manual(values = c("#E8E3D3", "#C4A055", "#85754D", "#4B2E83")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    title = "Washington Districts by Size (2025)",
    subtitle = "130 small districts serve fewer than 500 students each",
    x = "District Size",
    y = "Number of Districts"
  )
```

---

## 9. Tacoma held steady at ~30K for a decade before COVID

The state's third-largest city maintained remarkably stable enrollment from 2010 to 2020 before losing 2,000 students.

```{r tacoma-data}
tacoma <- enr |>
  filter(is_district, subgroup == "total_enrollment", grade_level == "TOTAL",
         grepl("Tacoma", district_name, ignore.case = TRUE)) |>
  select(end_year, district_name, n_students)
stopifnot(nrow(tacoma) > 0)

tacoma
```

```{r tacoma-chart}
print(tacoma)

ggplot(tacoma, aes(x = end_year, y = n_students)) +
  geom_line(linewidth = 1.2, color = "#4B2E83") +
  geom_point(size = 3, color = "#4B2E83") +
  geom_vline(xintercept = 2020.5, linetype = "dashed", color = "red", alpha = 0.5) +
  annotate("text", x = 2020.8, y = max(tacoma$n_students, na.rm = TRUE),
           label = "COVID", hjust = 0, color = "red", size = 3) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Tacoma School District Enrollment (2010-2025)",
    subtitle = "Remarkably stable for a decade, then COVID disruption",
    x = "School Year (ending)",
    y = "Total Enrollment"
  )
```

---

## 10. From 1.03M to 1.1M: 16 years of enrollment milestones

Washington's enrollment data spans 2010-2025, documenting the tech boom, pandemic disruption, and demographic transformation.

```{r summary-table}
decade_summary <- enr |>
  filter(is_state, subgroup == "total_enrollment", grade_level == "TOTAL",
         end_year %in% c(2010, 2015, 2019, 2021, 2025)) |>
  select(end_year, n_students) |>
  mutate(label = case_when(
    end_year == 2010 ~ "Post-recession",
    end_year == 2015 ~ "Tech boom",
    end_year == 2019 ~ "Pre-COVID peak",
    end_year == 2021 ~ "COVID low",
    end_year == 2025 ~ "Current"
  ))
stopifnot(nrow(decade_summary) > 0)

decade_summary
```

```{r summary-chart}
print(decade_summary)

ggplot(decade_summary, aes(x = end_year, y = n_students)) +
  geom_col(fill = "#4B2E83", width = 2) +
  geom_text(aes(label = paste0(label, "\n", scales::comma(n_students))),
            vjust = -0.3, size = 3) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Washington Enrollment Milestones",
    subtitle = "Key years from post-recession through COVID recovery",
    x = "School Year (ending)",
    y = "Total Enrollment"
  )
```

---

## 11. 1 in 6 Washington students now receives special education services

Special education identification rose from 13.2% to 16.4% over 16 years, with the rate accelerating since 2022.

```{r sped-data}
sped_trend <- enr |>
  filter(is_state, grade_level == "TOTAL", subgroup == "special_ed") |>
  select(end_year, n_students, pct) |>
  mutate(pct = round(pct * 100, 1))
stopifnot(nrow(sped_trend) > 0)

sped_trend
```

```{r sped-chart}
print(sped_trend)

ggplot(sped_trend, aes(x = end_year, y = pct)) +
  geom_line(linewidth = 1.2, color = "#4B2E83") +
  geom_point(size = 3, color = "#4B2E83") +
  geom_vline(xintercept = 2020.5, linetype = "dashed", color = "red", alpha = 0.5) +
  annotate("text", x = 2020.8, y = max(sped_trend$pct, na.rm = TRUE),
           label = "COVID", hjust = 0, color = "red", size = 3) +
  scale_y_continuous(labels = function(x) paste0(x, "%"),
                     limits = c(0, NA)) +
  labs(
    title = "Special Education Students as Share of Enrollment",
    subtitle = "Washington state, 2010-2025 -- rate accelerating since 2022",
    x = "School Year",
    y = "Percent of Students"
  )
```

---

## 12. Yakima Valley: where 93% of students are Hispanic and half are English learners

Central Washington's agricultural heartland has the state's highest concentrations of English learners and economically disadvantaged students.

```{r yakima-data}
yakima_districts <- enr |>
  filter(is_district, grade_level == "TOTAL", end_year == 2025,
         grepl("Yakima|Sunnyside|Toppenish|Wapato|Grandview", district_name)) |>
  select(district_name, subgroup, n_students) |>
  pivot_wider(names_from = subgroup, values_from = n_students) |>
  mutate(
    pct_hispanic = round(hispanic / total_enrollment * 100, 1),
    pct_ell = round(lep / total_enrollment * 100, 1),
    pct_econ_disadv = round(econ_disadv / total_enrollment * 100, 1)
  ) |>
  select(district_name, total_enrollment, pct_hispanic, pct_ell, pct_econ_disadv) |>
  arrange(desc(total_enrollment))
stopifnot(nrow(yakima_districts) > 0)

yakima_districts
```

```{r yakima-chart}
yakima_long <- enr |>
  filter(is_district, grade_level == "TOTAL", end_year == 2025,
         grepl("Yakima|Sunnyside|Toppenish|Wapato|Grandview", district_name),
         subgroup %in% c("hispanic", "lep", "econ_disadv")) |>
  select(district_name, subgroup, pct) |>
  mutate(pct = pct * 100,
         subgroup = case_when(
           subgroup == "hispanic" ~ "Hispanic",
           subgroup == "lep" ~ "English Learner",
           subgroup == "econ_disadv" ~ "Economically Disadvantaged"
         ))
stopifnot(nrow(yakima_long) > 0)

print(yakima_long)

ggplot(yakima_long, aes(x = reorder(district_name, -pct), y = pct, fill = subgroup)) +
  geom_col(position = "dodge") +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Yakima Valley Districts: Demographics",
    subtitle = "High concentrations of Hispanic, ELL, and low-income students",
    x = NULL,
    y = "Percent of Students",
    fill = "Subgroup"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
```

---

## 13. Ridgefield grew 84% while Vancouver lost 7% -- Clark County's diverging paths

The Portland metro spillover has reshaped Clark County, with small suburban districts booming while the urban core declines.

```{r clark-county-data}
clark_districts <- enr |>
  filter(is_district, subgroup == "total_enrollment", grade_level == "TOTAL",
         grepl("Vancouver|Evergreen SD|Camas|Battle Ground|Ridgefield", district_name),
         end_year %in% c(2015, 2020, 2025)) |>
  select(district_name, end_year, n_students) |>
  pivot_wider(names_from = end_year, values_from = n_students, names_prefix = "yr_") |>
  mutate(
    growth_2015_2025 = round((yr_2025 - yr_2015) / yr_2015 * 100, 1)
  ) |>
  arrange(desc(growth_2015_2025))
stopifnot(nrow(clark_districts) > 0)

clark_districts
```

```{r clark-county-chart}
clark_trend <- enr |>
  filter(is_district, subgroup == "total_enrollment", grade_level == "TOTAL",
         grepl("Vancouver|Evergreen SD|Camas|Battle Ground|Ridgefield", district_name))
stopifnot(nrow(clark_trend) > 0)

print(clark_trend |> filter(end_year == 2025) |> select(district_name, n_students))

ggplot(clark_trend, aes(x = end_year, y = n_students, color = district_name)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  geom_vline(xintercept = 2020.5, linetype = "dashed", color = "red", alpha = 0.5) +
  scale_y_continuous(labels = scales::comma) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Clark County District Enrollment Trends",
    subtitle = "Ridgefield booming while Vancouver declines",
    x = "School Year",
    y = "Enrollment",
    color = "District"
  ) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(nrow = 2))
```

---

## 14. Foster care enrollment dropped 53% from 2019 peak

Washington tracks foster care enrollment at every school, revealing the population peaked at 7,573 in 2019 and has since dropped to 3,560.

```{r foster-care-data}
foster_trend <- enr |>
  filter(is_state, grade_level == "TOTAL", subgroup == "foster_care") |>
  select(end_year, n_students, pct) |>
  mutate(pct = round(pct * 100, 2))
stopifnot(nrow(foster_trend) > 0)

foster_trend
```

```{r foster-care-chart}
print(foster_trend)

ggplot(foster_trend, aes(x = end_year, y = n_students)) +
  geom_line(linewidth = 1.2, color = "#4B2E83") +
  geom_point(size = 3, color = "#4B2E83") +
  geom_vline(xintercept = 2020.5, linetype = "dashed", color = "red", alpha = 0.5) +
  annotate("text", x = 2020.8, y = max(foster_trend$n_students, na.rm = TRUE),
           label = "COVID", hjust = 0, color = "red", size = 3) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Foster Care Students in Washington Schools",
    subtitle = "Peaked at 7,573 in 2019, now down to 3,560",
    x = "School Year (ending)",
    y = "Foster Care Students"
  )
```

---

## 15. The Tri-Cities boom: Pasco grew 31% while Kennewick and Richland added 19% and 32%

The Tri-Cities in southeastern Washington have seen sustained population growth driven by Hanford cleanup and agricultural expansion.

```{r tri-cities-data}
tri_cities <- enr |>
  filter(is_district, subgroup == "total_enrollment", grade_level == "TOTAL",
         grepl("Richland|Kennewick|Pasco", district_name)) |>
  select(end_year, district_name, n_students)
stopifnot(nrow(tri_cities) > 0)

tri_cities_wide <- tri_cities |>
  filter(end_year %in% c(2010, 2015, 2020, 2025)) |>
  pivot_wider(names_from = end_year, values_from = n_students, names_prefix = "yr_")

tri_cities_wide
```

```{r tri-cities-chart}
print(tri_cities_wide)

ggplot(tri_cities, aes(x = end_year, y = n_students, color = district_name)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  geom_vline(xintercept = 2020.5, linetype = "dashed", color = "red", alpha = 0.5) +
  scale_y_continuous(labels = scales::comma) +
  scale_color_manual(values = c("Kennewick School District" = "#4B2E83",
                                "Pasco School District" = "#C4A055",
                                "Richland School District" = "#85754D"),
                     labels = function(x) gsub(" School District", "", x)) +
  labs(
    title = "Tri-Cities Enrollment Growth",
    subtitle = "Southeastern Washington's sustained growth region",
    x = "School Year",
    y = "Enrollment",
    color = "District"
  ) +
  theme(legend.position = "bottom")
```

---

## Summary

Washington's school enrollment data reveals:

- **Steady growth**: 1.03M to 1.15M students over a decade before COVID
- **Seattle paradox**: The largest district lost 9% while suburban districts boomed
- **Demographic transformation**: White students dropped from 64% to 48%, Hispanic nearly doubled
- **Kindergarten cliff**: K enrollment down 14% from peak, signaling smaller future cohorts
- **Rural pressure**: 130 districts serve fewer than 500 students each
- **Special education growth**: Rising from 13% to 16% of all students
- **Agricultural communities**: Yakima Valley districts are 80-93% Hispanic
- **Clark County divergence**: Ridgefield up 84%, Vancouver down 7%
- **Foster care decline**: Peaked at 7,573 in 2019, now down 53%
- **Tri-Cities expansion**: All three districts grew 19-32% since 2010

These patterns shape school funding, facility planning, and staffing decisions across the Evergreen State.

---

*Data sourced from the Washington Office of Superintendent of Public Instruction (OSPI) via the [Washington State Report Card](https://washingtonstatereportcard.ospi.k12.wa.us/) and [data.wa.gov](https://data.wa.gov/).*

---

## Session Info

```{r session-info}
sessionInfo()
```
